---
title: "DATA 607 Week 3 Assignment - Strings and Regular Expressions"
author: "Alan Noel"
date: "9/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1 - Find "DATA" and "STATISTICS" 

Using the 173 majors listed in fivethirtyeight.com's College Majors data set 
provide code that identifies the majors that contain either
"DATA" or "STATISTICS".

The fivethirtyeight data set is at:

https://github.com/fivethirtyeight/data/blob/master/college-majors/majors-list.csv

## Download dataset CSV file from my GitHub repository. 


```{r retrieve-majors-list}
library(curl)
majors <- 
read.csv (curl("https://raw.githubusercontent.com/alnoel/CUNYSPS-Data607/main/majorslist.csv"))
head(majors)

```

## Create a Function to Find the Majors with the Search Words

```{r create-function}
findmajor <- function(srchword){
  
  loc <- grep(srchword,majors$Major)
  return(majors$Major[loc])
}
```

##  Place values we are looking for into a character vector
```{r create-list}
majorlist <- c("DATA","STATISTICS")
```

## Apply the function to all values in the character vector 
## and store values in a vector. Then loop through the vector
## and print out the majors found

```{r}
holdmajors <- sapply(majorlist, findmajor)
for (i in 1:length(holdmajors)){
  print(holdmajors[i])

}

```


# Task 2 - Take a List of Character Values in A Vector
# and Put Out the Creation Statement 
# for that Vector 

Create Vector

```{r create-vector}
raw.data <- c("bell pepper", "bilberry", "black berry",  "blood orange", 
              "blueberry", "cantaloupe", "chili pepper", "cloudberry",
              "elder berry", "lime lychee", "mulberry", "olive", "salal berry")
```


Convert the character vector of the terms to a single string
with commas separating the terms, just commas, no spaces 

```{r create-string}
hold <- paste(raw.data,sep = "", collapse = ",")
```

Output the vector creation statement a bit at a time using the cat function. 
Loop through the string looking for commas.At the point a comma is encountered 
we are the end of a value/term.


```{r find-terms}
itemcount <- 0



for (i in 1:nchar(hold)){
  if (i ==1){
    cat("c")
    cat('("')}

  # if not a comma then keep outputting
  # characters one at a time
  
    if (substr(hold,i,i) != ","){
    cat(substr(hold,i,i))
    
  }
  
  
  # if at a comma output one surrounded 
  # by double quotes. 
  
  # track items so at about halfway
  # through the set go to a new line
  # keep output tidy and within page width
  # break at a comma and not within an item
  
  if (substr(hold,i,i) == ","){
    cat('","')
    itemcount <- itemcount + 1
  }
  
  
  if ((itemcount == 6 ) & (substr(hold,i,i) == ",")){
    
    cat("\n")
    
  }
  if (i== nchar(hold)){
    cat('")')
  }
  
}



```


# Task 3 - Describe, in words, what these expressions will match:


(.)\1\1   This will not find anything. It will error. It is missing enclosure
by double quotes. Also We have one capturing group 
and it then refers to that capturing group twice but it does not
escape the tagging of the capturing group properly. It is escaped incorrectly 
using one backslash when it should use two. It's an attempt at 
"backreferencing". A capturing group stores the part of the string matched by
the part of the regular expression inside the parentheses.
I can run it against a vector of values with a couple that have three characters
in a row and it returns nothing. But with two backslashes it will find 
the same character appearing three times in a row. For example, "aaa". 
With two backslashes it is saying to use the capturing group twice. 

```{r}
library(stringr)
stuff <- c("111","123","aaa","zxy")
str_subset(stuff,"(.)\1\1")
str_subset(stuff,"(.)\\1\\1")
```


"(.)(.)\\\2\\\1"   This will find two characters followed by those characters
in reverse order. An example might be "cddc". It does this by invoking
the second capture group with the two and then the first capture group with
the 1 thus getting the reversal matching. 
   
```{r two-then-reverse}
stuff <- c("abdc","cddc","1234", "xyyx")
str_subset(stuff, "(.)(.)\\2\\1")


```
 

(..)\1  This will not work. It will error. It is missing enclosure
by double quotes.The use of backreferencing will not work 
because the backslash has not been repeated, escaped. It could be corrected to 
find a pair of characters repeated such as "klkl" or "abab123".
When I correct it then we get hits on repeating any pair of two characters
one right after the other.  


```{r two-pairs}
stuff <- c("klkl","abab123","123yoyo","ab1ab")
str_subset(stuff,"(..)\1" )
str_subset(stuff, "(..)\\1")
```


"(.).\\1.\\1" A character followed by any character, the original character,
any other character, the original character again. It is using the 
capture group labeled "1" two times so whatever matches the first time, 
which could be anything the first time, has two more times 
with any character interspersed. 
This will match, "yuyiy" and this will not "yuyix". 


```{r repeat-origina-3-times}
stuff <- c("abaca","12123", "12131")
str_subset(stuff, "(.).\\1.\\1")

```


"(.)(.)(.).*\\3\\2\\1"  Three characters followed by zero, or more 
characters, of any kind, followed by the original three characters
but in reverse order. 

```{r three-zero-reverse-three}
stuff <- c("abccba","abcxcba", "123x329","123329","123xx321")
str_subset(stuff, "(.)(.)(.).*\\3\\2\\1")

```


# Task 4 - Construct regular expressions to match words that:


## Start and end with the same character.



```{r start-end-same-character}

library(stringr)
stuff <- c("else","left","right","dead")
str_subset(stuff,"^(.)((.*\\1$)|\\1?$)")
```



## Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r repeated-pair-letters}

stuff <- c("cart","horse","church","mouse","ab888ab")
str_subset(stuff,"([A-Za-z][A-Za-z]).*\\1")
```



## Contain one letter repeated in at least three places (e.g. "eleven" 
## contains three "e"s.)

```{r one-letter-atleast-3places}
stuff <- c("eleven","available","rabbit","chevrolet")
str_subset(stuff,"([a-z]).*\\1.*\\1")


```






