---
title: "Assignment 1 - Basic Loading and Transformation:  \nUsing  \nFiveThiryEight Story 'Where Police Have Killed Americans in 2015'"
author: "Alan Noel"
date: "8/28/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This data is from the fivethirtyeight web site and documents police killings in the 
United states for the year 2015. It is the data behind the fivethirtyeight story
"Where Police Have Killed Americans in 2015". The data set was pulled
together using data collected by the Guardian newspaper of the UK and  
from the US Census. 

My purpose in using this data set was to investigate how 
many blacks and black males are the victims. Are the amounts disproportionate?  

# Get Data and Do Some Exploration

## Retrieve Raw CSV file from GitHub

I placed a copy of fivethirtyeight's data set onto my area of 
GitHub. 

```{r retrieve csv file from GitHub}
library(curl)
origcsv <- 
read.csv( curl("https://raw.githubusercontent.com/alnoel/CUNYSPS-Data607/main/police_killings.csv"))
```

## Display Structure of the Data Frame Created from the Original CSV File 

The data set comprises 467 observations of 38 variables. I find it 
interesting that none of the character variables imported as factors
even though I did not specify "stringsAsFactors = FALSE". Although
factors are useful in analysis my limited experience has been to 
prefer to make factors as I need them and not get them automatically.

```{r structure}
str(origcsv)
```

# Dataframe Adjustments

## Drop Some of the Columns Not of Interest to My Purposes 

```{r drop columns}
# drop columns not of interest at the moment

df_trimmed = 
subset(origcsv, select = -c(state_fp,county_fp, tract_ce,geo_id,  
                            share_white,share_black,share_hispanic))
```

## Rename Some of the Columns to be More Meaningful

```{r rename columns}
library(dplyr)
df_trimmed <- rename(df_trimmed,
                     Tract_description = namelsad,
                     Tract_unemployment = urate,
                     Tract_population = pop
                     )

```

# Analysis 

## What are the proportions by racial/ethnic groups? 

```{r}

round(prop.table(table(df_trimmed$raceethnicity)),2)
```

It is easier to see proportions with a bar plot

```{r}
counts <- prop.table(table(df_trimmed$raceethnicity))
barplot(counts, main ="Proportions By Race 2015", cex.names = 0.6, xlab ="Race",
        ylab = "Proportion",ylim = c(0.0,1.0))
abline(h = 0.5)
```

17.9 percent of the US Population was black in 2015 yet 29% of
those killed by police were black. 76.3% of the population was white
but only 51% of those killed by police were white. 


## What is the proportion that is black males? 

```{r NA and Calculate Proportion}
# change all race ethnicity values to NA if "Unknown"

df_trimmed$raceethnicity <- na_if(df_trimmed$raceethnicity,"Unknown")

# Determine the number of black males

num_blackmale <- sum(df_trimmed$raceethnicity == "Black" & df_trimmed$gender =="Male",na.rm=TRUE)
num_obs  <-nrow(df_trimmed) 
(prcnt_blackmale <- num_blackmale / num_obs)

```

The majority of black victims are males. 29% of the
victims are black and 27% are black males. 


# Conclusion

There appears to be evidence of blacks and black males
being disproportionately the victims of police killings
in the US during the calendar year 2015. If this 
was a situation where we conducted a hypothesis test
I suspect we would be rejecting the null hypothesis
and concluding there is something other than
chance shaping the outcome.






