---
title: "SQL and R Assignment - Create Data in a DBMS and Get It to R"
author: "Alan Noel"
date: "9/2/2021"
output:
  html_document: default
  pdf_document: default
---

# Introduction
This exercise was to learn how to create some data in a 
DBMS, in this case MySQL, and get that data into R. 
The MySQL database joins together a couple of tables
of data regarding movie reviews. It then places that flattened set of data 
out into a CSV file. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load tidyverse


```{r tidy, warning = FALSE, message=FALSE}
library(tidyverse)
```

# Retrieve Data from GitHub

```{r load csv, echo=FALSE}
library(curl)
# moviereviews <- read_csv("moviereviews.csv")
moviereviews <- 
read.csv( curl("https://raw.githubusercontent.com/alnoel/CUNYSPS-Data607/main/moviereviews.csv"))

```

# Check for NA's in All the Columns - One Column, 'rating' Has Some NA's

```{r}
df <- as.data.frame(
  cbind(
    lapply(
      lapply(moviereviews, is.na), sum)
  )
)

rownames(subset(df, df$V1 != 0))
```

# Histogram of Ratings Before Imputation
```{r before imputation}
hist(moviereviews$rating, xlab= 'Ratings ', main = "Before Imputation\n Distribution of Ratings",col="darkmagenta")

```






# Use Median Imputation to Replace NA's in 'Rating' Column 

```{r impute missing values}
moviereviews <- moviereviews %>% 
                mutate(rating = ifelse(is.na(rating),
                                as.integer(median(rating, na.rm = TRUE)),
                              rating))
```

# Check Again to Ensure All NA's Replaced 

```{r impute check }
df <- as.data.frame(
  cbind(
    lapply(
      lapply(moviereviews, is.na), sum)
  )
)

rownames(subset(df, df$V1 != 0))
```


# Histogram of Ratings After Imputation
```{r after imputation}
hist(moviereviews$rating, xlab= 'Ratings ', main = "After Imputation\nDistribution of Ratings",col="darkmagenta")

```

# Observation

Imputation seems to provide values consistent with what I would expect. Because
all six movies are quite popular I would expect the ratings to tend high and 
the imputation results in a bit higher scores. I don't think it is the case 
the imputation biased the data but actually helped provide a more realistic 
data set. 


