---
title: "Project 1"
author: "Alan Noel"
date: "9/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Convert Chess Tournament Text File to CSV with Transformation

## Load Libraries Needed 

```{r load-libraries}
library(dplyr)
library(curl)
```
## Load Text File from GitHub
```{r load-text-file, warning=FALSE}
# Load text file into a character vector

hold <- readLines (curl("https://raw.githubusercontent.com/alnoel/CUNYSPS-Data607/main/tournamentinfo.txt"))
```

## Create Empty Data Frame to Hold Values Extracted From The Text File  

```{r ure, echo=FALSE}
extracts <- data.frame(pairnum=character(),
                       state=character(),
                       name=character(),
                       totalpoints=character(),
                       rnd1=character(),
                       rnd2=character(),
                       rnd3=character(),
                       rnd4=character(),
                       rnd5=character(),
                       rnd6=character(),
                       rnd7=character(),
                       prerating=character(),
                       stringsAsFactors=FALSE)
```

## Extract values 

Start at 5th item in the vector 'hold'. Each player has two items/rows in the vector.
Process the first row for the player then skip to second row for the player. Then
process the second row for the player. Skip to the next player and move values picked out
over to the "extracts" data frame to hold them for later determining the average 
pre chess rating of opponents. 


```{r loop-and-extract}

i <- 5
while (i < length(hold)){

  pairnum <- substr(hold[i],1,5)
  name <-  substr(hold[i],9,39)
  totalpoints <-  substr(hold[i],42,45)
  rnd1 <-  substr(hold[i],51,52)
  rnd2 <-  substr(hold[i],57,58)
  rnd3 <-  substr(hold[i],63,64)
  rnd4 <-  substr(hold[i],69,70)
  rnd5 <-  substr(hold[i],75,76)
  rnd6 <-  substr(hold[i],81,82)
  rnd7 <-  substr(hold[i],87,88)

  i = i +1
  state <- substr(hold[i],4,5)
  prerating <- substr(hold[i],23,29)
  extracts[nrow(extracts)+1,] = c(pairnum,state,name,totalpoints,rnd1,rnd2,rnd3,rnd4, rnd5,rnd6,
                                  rnd7,prerating)
  
  i = i + 2

}

```

## Remove All Before "P" in Prerating
```{r removeP}

extracts$prerating <- substr(extracts$prerating, 1,4)
```


## Replace All Instances of Two spaces, "  ", with NA
## in all 7 of the rnd's Columns

```{r two-spaces-to-NA, warning=FALSE}
for (i in 5:11){
  
  extracts[,i] <- gsub("  ",NA, extracts[,i])
}
```


## Convert the Following Variables/Columns to Numeric: 
# pairnum, totalpoints, rnd1-rnd7, prerating

```{r convert-numeric}

extracts$pairnum <- as.numeric(extracts$pairnum)
extracts$totalpoints <- as.numeric(extracts$totalpoints)
extracts$rnd1 <- as.numeric(extracts$rnd1)
extracts$rnd2 <- as.numeric(extracts$rnd2)
extracts$rnd3 <- as.numeric(extracts$rnd3)
extracts$rnd4 <- as.numeric(extracts$rnd4)
extracts$rnd5 <- as.numeric(extracts$rnd5)
extracts$rnd6 <- as.numeric(extracts$rnd6)
extracts$rnd7 <- as.numeric(extracts$rnd7)
extracts$prerating <- as.numeric(extracts$prerating)

```

## Create Empty Data Frame to Hold Final Results, a CSV File

```{r empty-dataframe}
finalresults <- data.frame(name=character(),
                           state=character(),
                           totalpoints=double(),
                           prerating=double(),
                           oppavgprerating=double(),
                           stringsAsFactors=FALSE)
```

## Calculate Average Pre Chess Rating of Opponents

```{r calculate-average-prechess}
for (i in 1:nrow(extracts)){
  calcvector <- vector()
  for (j in 5:11){
    if (!is.na(extracts[i,j])){
      
      oppnum <-extracts[i,j]
      opppre <- extracts[oppnum,12]
      calcvector <- c(calcvector,opppre)
    }   
  }
  avgprechessop <-round(mean(calcvector),0)

  
  finalresults[nrow(finalresults) +1,] =
    c(extracts[i,3], # name
      extracts[i,2], # state
      extracts[i,4], # total points
      extracts[i,12], # prerating
      avgprechessop  # avg pre chess rating of opponents
      
    )
}


```

## Use trimws to Remove leading and Trailing blank characters

```{r remove-leading-trailing}
finalresults$name <- trimws(finalresults$name)
```

## Adjust Total Points to Include Decimal Points in all Cases, i.e. 6.0, not 6
```{r adjust-total-points}
finalresults$totalpoints <- format(round(as.numeric(finalresults$totalpoints),2),nsmall=1)

```


## Write Data Frame to csv file. Don't include row names, i.e. "1","2","3", etc.
```{r write-csv-file }
write.csv(finalresults, "finalresults.csv",row.names= FALSE)

```

 


